---
title: "Pilot CaTS-RNA summary to share"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE, include=FALSE}

rm(list=ls())

library(DESeq2)
library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(tximport)

```


```{r, include=FALSE}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="inputs_RNApilot/"

outdir="outputs_glialRNA_1/"
ifelse(!dir.exists(file.path(workingdir,outdir)), dir.create(file.path(workingdir,outdir)), "Directory exists")

```


```{r, include=FALSE }


#salmon counts from pipeline, import
path_files =  list.files(paste0(workingdir,subworkinput)) 
samples = data.frame(run=path_files, stringsAsFactors = FALSE) %>%
  filter(str_detect(run, "^D"))

files <- file.path(paste0(workingdir,subworkinput), samples$run, "quant.sf")
names(files) <- samples$run
all(file.exists(files))

#from pipeline
tx2gene = read_tsv(paste0(paste0(workingdir,subworkinput),"/salmon_tx2gene.tsv"))

txi.salmon <- tximport(files, type = "salmon", tx2gene = tx2gene)
#head(txi.salmon$counts)

```


```{r, include=FALSE}

count.table <- txi.salmon$counts

```


```{r , include=FALSE}
sorted_gate <- c("pMN","DP","p3","neur")
sorted_conditions <- c("500","UPSAG","dRA2UPSAG","dRA")

# sorted.DayGate <- c("D5_p1","D5_p2","D5_pM",
#                     "D7_p1","D7_p2","D7_pM",
#                     "D9_p1","D9_p2","D9_pM",
#                     "D11_p1","D11_p2","D11_pM")
# sorted.day <- c("D5","D7","D9","D11")
# sorted.dayNfia <- c("D5_NFIAn_WT","D5_NFIAn_MUT",
#                     "D7_NFIAn_WT","D7_NFIAn_MUT",
#                     "D9_NFIAn_WT","D9_NFIAp_WT","D9_NFIAn_MUT",
#                     "D11_NFIAp_WT","D11_NFIAn_MUT")
# 

# 
# nfiaStatus_order <- c("NFIAn_WT","NFIAp_WT","NFIAn_MUT")

shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
# 
# # for Days
# red_colors <- c("#fadede","#f3aaaa","#e96666","#cf1e1e")


color_gates <- c("#b30000","#800080","#009640","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")

# conditions
colors_conditions <- c("#b81bb8","#b81b1b","#1bb81b","#1bb8b8")

```



```{r, include=FALSE}

count.matrix <- count.table

## Make metadata file for DESeq

genecolData_first <- data.frame(Sample_ID = colnames(count.matrix))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))


dds <- DESeqDataSetFromTximport(txi = txi.salmon,
                              colData = genecolData_first,
                              design = Condition ~ Gate)

dds <- DESeq(dds)

vsd <- varianceStabilizingTransformation(dds,blind = FALSE)

# Export normalized tables for plotting elsewhere
dds_counts <- counts(dds, normalized = TRUE)
vsd_data <- assay(vsd)

```

## Plot PCAs 

```{r , include=FALSE}
## to get more than 2 components

## to go it in the top ntop variable genes 
# calculate the variance for each gene
rv <- rowVars(vsd_data)
# select the ntop genes by variance
ntop=500
select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]

#pca <- prcomp(t(assay(object)[select,]))
t_vsd <- t(vsd_data[select,])
vsd_pca <- prcomp(t_vsd, retx=TRUE, scale. = FALSE)

names(vsd_pca)

# How many PC to explain enough variance?
#summary(vsd_pca)

var_explained <-vsd_pca$sdev^2/sum(vsd_pca$sdev^2)
plot(var_explained)

```

```{r, include=FALSE}

vsd_pca_plot <- vsd_pca$x %>% 
  as.data.frame %>%
  rownames_to_column("Sample") %>%
  separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Experiment=paste(Condition,Rep,sep="_"),
         Gate=factor(Gate, levels=sorted_gate))

```

#### PC1 is dominated by neurons vs progenitor
```{r }
ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Day)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```

#### Compared to that, even the temporal transition is quite minimal
```{r}

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Day,shape=Gate)) +
  scale_fill_manual(values = colors_greys) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```


#### PC2 seems to be driven by the sorted Gates

- DP nicely positioned between pMN and p3. Which condition they come from does not seem as critical, but they are colored in the next plot. 

```{r}

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```

#### Colored by condition

Maybe downRA is a bit more different than the rest. 

```{r}

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Condition,shape=Gate)) +
  scale_fill_manual(values = colors_conditions) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```

#### Maybe (only maybe) DP at later timepoints a bit more similar to pMN

This is PCA so distances does matter but still, further analysis needed. 

```{r}
ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Day)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)
```



There might be some outliers in the other PCs but as you can see it'll be very little variance. 


### Correlation Heatmaps - based on 500 top genes 

Mainly cluster by sorted gate

```{r fig.width=11,fig.height=11, include=FALSE}

cor_vsd <-cor( vsd_data[select,] )

names(color_gates) <- sorted_gate
names(colors_conditions) <- sorted_conditions

phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Condition","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, Condition=colors_conditions,
    Day = c(D5="#f6f6f6",D6="#808080",D7="#333333"),
  Rep = c(R1="#ebeb77", R2="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)

```


```{r fig.width=11,fig.height=11 }
Heatmap(cor_vsd, name="Correlation", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Pearson Correlation - Top 500 genes")



```




```{r, include=FALSE}
# Export normalized tables for plotting elsewhere
dds_counts <- counts(dds, normalized = TRUE)
vsd_data <- assay(vsd)


```

#### Looking at some key genes: colored by condition

```{r, include=FALSE}

dds_counts_plot <- dds_counts %>% 
  as.data.frame() %>%
  rownames_to_column("geneid") %>%
  gather(sampleid, counts_norm, starts_with("D")) %>%
  separate(sampleid,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Gate=factor(Gate, levels=sorted_gate),
         Condition=factor(Condition, levels=sorted_conditions))
```


- p3 from dRA express floor plate markes: expected by reassuring. 

- The rest of progenitors: express markers of their gate, regarless of condition of origin (500, upSAG, dRA2UPSAG)

- Neurons: UpSAG and 500 SAG are making Motor Neurons. What neurons are the other conditions making?

- The exception is Dbx1 but remember that comes from having been in 0-SAG for 1 day and we have seen those pMN take 1 day to lower the levels of Pax6 (by flow). Dbx1 expression is consistent with their transitioning from p0. 

```{r fig.height=9, fig.width=7}

geneOI <- c("Sox2","Dbx1","Irx3","Pax6","Nkx6-1","Olig2","Nkx2-2","Foxa2","Shh","Arx","Tubb3","Sim1","Mnx1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free") +
  theme_bw()

```


#### Now colored by gate

To spot differences in conditions or days. 

- Apart from p3 from dRA being floor plate, I don't see anything major. 

```{r fig.height=9, fig.width=9}
ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```

#### What next?

- Differential expression: for conditions and gates with 2 replicates
- Interrogating other genes (i.e. the ones from Hynek's human DP)


```{r, include=FALSE}
sessionInfo()
```


