---
title: "BARRY_2_230919_run_H1 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}

rm(list=ls())

library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(UpSetR)

```

# BARRY analysis

First attempt with aborted run. 

On 5th Dec 2023 after hamming distance of 1.

### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="input_BARRY_2/H1_counts/"

```


## Colors and shapes
Before more plotting, let's get some metadata organised

```{r }
sorted_gate <- c("pMN","DP","p3","DN")

shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
 

color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")


```

### import data

```{r  import-domain}

barry_data <- lapply(list.files(path=paste0(workingdir,subworkinput),pattern=".*.fq.gz.txt", full.names=TRUE),function(x) {
  data <- read.csv(x,header=F,stringsAsFactors=F) %>% as.data.frame()
  data$sample <- gsub(paste0("/Users/delasj/Dropbox \\(The Francis Crick\\)/DP_cisReg/",subworkinput,"/"),"", x)
  data$sample <- gsub("_R1_001_trimmed.fq.gz.txt","",data$sample)
  #data <- data %>% column_to_rownames(sample)
  data
})

barry_results <- do.call(rbind,barry_data) %>% column_to_rownames("sample")

```


#### Convert any count > 0 to 1
Make binary matrix for quick comparisons

```{r}
# test1 <- barry_results
# test1 <- test1 %>% column_to_rownames("sample")
# test <- test1[1:3,5:15]

barry_binary <- as.matrix((barry_results>0)+0)

```

#### Some numbers

```{r, fig.width=8, fig.height=4}

barcodes_per_sample <- rowSums(barry_binary) %>% as.data.frame() %>%
  rownames_to_column("sampleid") %>%
  separate(sampleid, into = c("Day","Gate","Rep","discard"), sep = "_", remove = FALSE) 


ggplot(barcodes_per_sample, aes(x=Gate,y=get(names(barcodes_per_sample)[6]))) +
  geom_col(aes(fill=Rep), position=position_dodge()) +
  facet_grid(~ Day) +
  ylab("Unique barcodes per sample") +
  theme_bw()


```



```{r}

#test_binary <- barry_binary[1:4,5:35]

barcodes_shared <- colSums(barry_binary) %>% as.data.frame() %>% 
  rownames_to_column("BCid")

colnames(barcodes_shared) <- c("BCid","shared_samples")

ggplot(barcodes_shared, aes(x=reorder(BCid,-shared_samples), y=shared_samples)) +
  geom_col() + 
  theme_minimal() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))

```

Product of matrix by t(matrix)


```{r}

i <- (colSums(barry_binary, na.rm=T) > 1) # T if colSum > 1, F otherwise
binary_keepshared <- barry_binary[, i] # all the non-zero columns

binary_keepshared_matrix2 <- binary_keepshared%*%t(binary_keepshared)

# binary_keep_samplename <- binary_keepshared_matrix2[,5:20] %>% as.data.frame() %>%
#   rownames_to_column("sampleid") %>%
#   separate(sampleid, into = c("Day","Gate","Rep","discard"), sep = "_", remove = FALSE)



```


```{r, fig.width=12, fig.height=10}

names(color_gates) <- sorted_gate


genecolData_first <- data.frame(Sample_ID = colnames(binary_keepshared_matrix2))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, 
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)


Heatmap(binary_keepshared_matrix2, name="Shared barcodes", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Shared number of barcodes")




```


## For now repeat this with separeted replicates

```{r}



binary_keep_repA <- binary_keepshared %>% as.data.frame() %>%
  rownames_to_column("sampleid") %>%
  separate(sampleid, into = c("Day","Gate","Rep","discard"), sep = "_", remove = FALSE) %>%
  filter(Rep=="A") %>%
  select(!c("Day","Gate","Rep","discard")) %>%
  column_to_rownames("sampleid") %>%
  as.matrix()

binary_keep_repB <- binary_keepshared %>% as.data.frame() %>%
  rownames_to_column("sampleid") %>%
  separate(sampleid, into = c("Day","Gate","Rep","discard"), sep = "_", remove = FALSE) %>%
  filter(Rep=="B") %>%
  select(!c("Day","Gate","Rep","discard")) %>%
  column_to_rownames("sampleid") %>%
  as.matrix()



binary_keepshared_matrixA <- binary_keep_repA%*%t(binary_keep_repA)
binary_keepshared_matrixB <- binary_keep_repB%*%t(binary_keep_repB)

```

```{r, fig.width=8, fig.height=7}

names(color_gates) <- sorted_gate


genecolData_first <- data.frame(Sample_ID = colnames(binary_keepshared_matrixA))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates,
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)


Heatmap(binary_keepshared_matrixA, name="Shared barcodes", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Shared number of barcodes")




```



```{r, fig.width=8, fig.height=7}

names(color_gates) <- sorted_gate



genecolData_first <- data.frame(Sample_ID = colnames(binary_keepshared_matrixB))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, 
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)


Heatmap(binary_keepshared_matrixB, name="Shared barcodes", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Shared number of barcodes")




```

### just to try
remove barcodes shared across many populations


```{r}

i <- (colSums(barry_binary, na.rm=T) > 1 & colSums(barry_binary, na.rm=T) < 6 ) # T if colSum > 1, F otherwise
binary_keepshared <- barry_binary[, i] # all the non-zero columns

binary_keepshared_matrix2 <- binary_keepshared%*%t(binary_keepshared)

```


```{r, fig.width=12, fig.height=10}

names(color_gates) <- sorted_gate


genecolData_first <- data.frame(Sample_ID = colnames(binary_keepshared_matrix2))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, 
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)


Heatmap(binary_keepshared_matrix2, name="Shared barcodes", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Shared number of barcodes")




```
## Option 2
### counts per barcode, then heatmap

```{r}

barry_results_t <- t(barry_results)

barry_results_countsperbc <- rowSums(barry_results_t) %>% as.data.frame() %>%
  rownames_to_column("BCid")

colnames(barry_results_countsperbc) <- c("BCid","row_sumcounts")

ggplot(barry_results_countsperbc, aes(x=reorder(BCid,-row_sumcounts), y=row_sumcounts)) +
  geom_col() + 
  theme_minimal() +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))


```

```{r}

dim(barry_results_countsperbc)

dim(barcodes_shared)

counts_vs_shared <- barry_results_countsperbc %>% left_join(barcodes_shared, by = "BCid")

ggplot(counts_vs_shared, aes(x=row_sumcounts, y=shared_samples)) +
  geom_point(alpha=0.5) + 
  theme_minimal()

ggplot(counts_vs_shared, aes(x=row_sumcounts, y=shared_samples)) +
  geom_point(alpha=0.1, position=position_jitter(width=0, height = 0.25)) + 
  theme_minimal()

ggplot(counts_vs_shared, aes(x=row_sumcounts, y=shared_samples)) +
  geom_hex(bins = 70) +
  scale_fill_continuous(type = "viridis") +
  theme_bw()

```

### possibly filter?


above sumcounts of 5000
and below 1250


This was 8677 barcodes

We also need to filter barcodes only in 1 sample: left_join

Now it's 8668

```{r}

#head(barry_results_t)

barry_results_t_filter <- barry_results_t %>%
  as.data.frame() %>%
  filter(rowSums(.) <5000 & rowSums(.) > 1000) %>%
  rownames_to_column("BCid") %>%
  left_join(barcodes_shared, by = "BCid") %>% # this is to filter out single appearance bcs.
  filter(shared_samples > 1) %>%
  select(-shared_samples) %>%
  column_to_rownames("BCid") %>%
  as.matrix()


dim(barry_results_t_filter)

### sadly, if I want to plot per rep I have to clean up per rep

## min counts per rep
barry_results_t_A <- barry_results_t %>% as.data.frame %>% select(matches("_A_"))
barry_results_countsperbc <- rowSums(barry_results_t_A) %>% as.data.frame() %>%
  rownames_to_column("BCid")

colnames(barry_results_countsperbc) <- c("BCid","row_sumcounts")

ggplot(barry_results_countsperbc, aes(x=reorder(BCid,-row_sumcounts), y=row_sumcounts)) +
  geom_col() + 
  theme_minimal() +
  geom_hline(yintercept=c(500,5000), color="red", linetype="dashed")+
  ggtitle("repA") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))


barry_results_t_B <- barry_results_t %>% as.data.frame %>% select(matches("_B_"))
barry_results_countsperbc <- rowSums(barry_results_t_B) %>% as.data.frame() %>%
  rownames_to_column("BCid")

colnames(barry_results_countsperbc) <- c("BCid","row_sumcounts")

ggplot(barry_results_countsperbc, aes(x=reorder(BCid,-row_sumcounts), y=row_sumcounts)) +
  geom_col() + 
  theme_minimal() +
  geom_hline(yintercept=c(500,5000), color="red", linetype="dashed")+
  ggtitle("repB") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))



barry_results_t_filterA <- barry_results_t %>%
  as.data.frame() %>%
  rownames_to_column("BCid") %>%
  left_join(barcodes_shared, by = "BCid") %>% # this is to filter out single appearance bcs.
  filter(shared_samples > 1) %>%
  select(-shared_samples) %>%
  column_to_rownames("BCid") %>%
  select(matches("_A_")) %>% #separate A and B
  filter(rowSums(.) <5000 & rowSums(.) > 500) %>% 
  as.matrix()

barry_results_t_filterB <- barry_results_t %>%
  as.data.frame() %>%
  rownames_to_column("BCid") %>%
  left_join(barcodes_shared, by = "BCid") %>% # this is to filter out single appearance bcs.
  filter(shared_samples > 1) %>%
  select(-shared_samples) %>%
  column_to_rownames("BCid") %>%
  select(matches("_B_")) %>% #separate A and B
  filter(rowSums(.) <5000 & rowSums(.) > 500) %>%
  as.matrix()

dim(barry_results_t_filterA)
dim(barry_results_t_filterB)

```




### heatmap?

RepA

```{r}


genecolData_first <- data.frame(Sample_ID = colnames(barry_results_t_filterA))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, 
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(brewer.pal(n = 9, name = "Reds"))(100)

```

```{r}

heatmapA <- Heatmap(barry_results_t_filterA, name="Barcode counts repA", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        show_row_names = FALSE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        top_annotation = heatmap_ann,
        column_title = "Barcode counts")




```


RepB

```{r}


genecolData_first <- data.frame(Sample_ID = colnames(barry_results_t_filterB))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Gate","Rep","discard"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))



phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, 
    Day = c(D4="#f6f6f6",D5="#cecece",D6="#808080",D7="#333333"),
  Rep = c(A="#ebeb77", B="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(brewer.pal(n = 9, name = "Reds"))(100)

```

```{r fig.width=9, fig.height=9}

heatmapB <- Heatmap(barry_results_t_filterB, name="Barcode counts rep B", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        show_row_names = FALSE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        top_annotation = heatmap_ann,
        column_title = "Barcode counts")




```

```{r fig.width=9, fig.height=9}

draw(heatmapA)
draw(heatmapB)
```



```{r}
sessionInfo()
```
