---
title: "Segmentation analysis FOXA2 cKO - genotypes"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}

rm(list=ls())

library(RColorBrewer)
library(tidyverse)


```

# Segmentation analysis

This is from the table generated by Cell Profiler

There were incorrectly assigned animals. Now we have a lookup table. 

### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="inputs_foxa2KOsegment_1/"

```


## Colors and shapes
Before more plotting, let's get some metadata organised

```{r }
sorted_gate <- c("pMN","DP","p3","DN")

shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
 

color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")


```

### import data

```{r  import-segment}

segment_cp <- read.csv(file=paste0(workingdir,subworkinput,"Joaquina__Dec_6_2023masks_objects.csv"),header=TRUE, stringsAsFactors = FALSE)
#colnames(segment_cp)

segment_cp_clean <- segment_cp %>% select(c("ImageNumber","FileName_DAPI_Image","Intensity_MeanIntensity_DAPI_Image",
                      "Intensity_MeanIntensity_FOXA2_Image","Intensity_MeanIntensity_NKX22_Image",
                      "Intensity_MeanIntensity_OLIG2_Image","Intensity_MeanIntensity_SOX2_Image",
                      "Location_Center_X","Location_Center_Y"))  %>%
                    rename("embryoID"="FileName_DAPI_Image") %>%
                    mutate(embryoID=gsub("C1-","",embryoID)) %>%
                    mutate(embryoID=gsub(".tif","",embryoID))



```

```{r import-lookup}

lookup_id2geno <- read.csv(file=paste0(workingdir,subworkinput,"lookup_embryoID_genotype.csv"),header=TRUE, stringsAsFactors = FALSE)

```



## Filter embryos for which the table has genotypes

and edit names 

```{r}

segment_cp_geno <- segment_cp_clean %>%
  inner_join(lookup_id2geno, by="embryoID") %>%
  filter(embryoWellid != "") %>%
  separate(embryoID, into =c("block","position","embryonumber"), sep="_") %>%
  mutate(embryoID = paste(block,position,embryoWellid, sep="_"))


```



#### Has any of this worked?


```{r}

ggplot(segment_cp_geno %>% filter(ImageNumber=="4"), aes(x=Location_Center_X,y=-Location_Center_Y)) +
  geom_point() +
  geom_point(aes(color=Intensity_MeanIntensity_FOXA2_Image))+
  ggtitle(segment_cp_geno[segment_cp_geno$ImageNumber=="4","embryoID"]) +
  coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

```

```{r fig.width=14, fig.height=12}

ggplot(segment_cp_geno, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_FOXA2_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_SOX2_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

```

## Find thresholds

```{r fig.width=12, fig.height=12}
sox2_thres = 0.25
nkx22_thress = 0.35
olig2_thress = 0.2
foxa2_thress = 0.3

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_SOX2_Image, y=Intensity_MeanIntensity_OLIG2_Image)) +
  geom_hex(binwidth = c(0.04, 0.04))  +
  geom_vline(xintercept=sox2_thres,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=olig2_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_SOX2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_hex(binwidth = c(0.04, 0.04))  +
  geom_vline(xintercept=sox2_thres,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_OLIG2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_hex(binwidth = c(0.04, 0.04))  +
  geom_vline(xintercept=olig2_thress,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_FOXA2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_hex(binwidth = c(0.04, 0.04))  +
  geom_vline(xintercept=foxa2_thress,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

```

```{r fig.width=14, fig.height=12}

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_FOXA2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_point(size=0.2, aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_vline(xintercept=foxa2_thress,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_OLIG2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_point(size=0.2, aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  geom_vline(xintercept=olig2_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_SOX2_Image, y=Intensity_MeanIntensity_NKX22_Image)) +
  geom_point(size=0.2, aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_vline(xintercept=sox2_thres,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(segment_cp_geno, aes(x=Intensity_MeanIntensity_FOXA2_Image, y=Intensity_MeanIntensity_SOX2_Image)) +
  geom_point(size=0.2, aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_vline(xintercept=foxa2_thress,  linetype="dashed", color = "blue") +
  geom_hline(yintercept=sox2_thres,  linetype="dashed", color = "blue") +
  # scale_x_log10()+
  # scale_y_log10() +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()


```

### Steps

- Filter out gut:
  1. by levels of SOX2
  2. by requiring SOX2 and 1 other channel
  3. WORKED: by position with respect to the pMN population
  
- Plot classify
- Plot number of Foxa2 cells vs DP


Classification based on Intensity_MeanIntensity_ANTIBODY_Image and the thresholds above.

sox2_thres = 0.25
nkx22_thress = 0.35
olig2_thress = 0.2
foxa2_thress = 0.3


```{r fig.width=12, fig.height=12}

full_classified <- segment_cp_geno %>%
  mutate(celltype=if_else(Intensity_MeanIntensity_OLIG2_Image > olig2_thress & Intensity_MeanIntensity_NKX22_Image > nkx22_thress, "DP",
                          if_else(Intensity_MeanIntensity_NKX22_Image > nkx22_thress, "p3",
                                  if_else(Intensity_MeanIntensity_OLIG2_Image > olig2_thress, "pMN",
                                      if_else(Intensity_MeanIntensity_SOX2_Image > sox2_thres,"Sox2_notp3","othercell"))))) %>%
  mutate(foxa2_positive=if_else(Intensity_MeanIntensity_FOXA2_Image>foxa2_thress, "Foxa2","notFoxa2"))

ggplot(full_classified, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.05, aes(color=celltype))+
  facet_wrap( ~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw() +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(full_classified, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.05, aes(color=foxa2_positive))+
  facet_wrap( ~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line = element_line(colour = "black"))

# ggplot(full_classified, aes(x=Location_Center_X,y=-Location_Center_Y)) +
#   #geom_point(size=0.01) +
#   geom_point(size = 0.3, stroke = 0, aes(color=celltype))+
#   facet_wrap( ~ embryoID) +
#   coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
#   theme_bw(base_size = 10) +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#   panel.background = element_blank(), axis.line = element_line(colour = "black"))
# 
# ggplot(full_classified, aes(x=Location_Center_X,y=-Location_Center_Y)) +
#   #geom_point(size=0.01) +
#   geom_point(size = 0.3, stroke = 0, aes(color=foxa2_positive))+
#   facet_wrap( ~ embryoID) +
#   coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
#   theme_bw(base_size = 10)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#   panel.background = element_blank(), axis.line = element_line(colour = "black"))



```

#### Filter background SOX2+ FOXA2+ cells based on position

Cell that are positioned more than 5 standard deviations away from the mean of the pMN population (dashed red line) will be excluded. Visualize the excluded cells below.


```{r fig.width=14, fig.height=12}

full_classified_mean_pmnposition <- full_classified %>%
  filter(celltype %in% c("pMN")) %>%
  group_by(ImageNumber) %>%
  mutate(ave_y = mean(Location_Center_Y, na.rm = TRUE),
         sd_y = sd(Location_Center_Y, na.rm = TRUE))

dotsize=0.2


# ggplot(full_classified_filter_gut, aes(x=Location_Center_X, y=-Location_Center_Y)) +
#   geom_point(size=dotsize,aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
#   scale_color_continuous(type = "viridis") +
#   expand_limits(y=c(0,-1024),x=c(0,1024)) +
#   geom_hline(data=full_classified_filter_gut, aes(yintercept=-(ave_y+2.5*sd_y)),linetype='dashed', col = 'red') +
#   facet_wrap(~ embryoID) +
#   coord_fixed(ratio = 1) +
#   theme_bw(base_size = 8)

ggplot(full_classified, aes(x=Location_Center_X, y=-Location_Center_Y)) +
  geom_point(size=dotsize,aes(color=Intensity_MeanIntensity_FOXA2_Image)) +
  scale_color_continuous(type = "viridis") +
  expand_limits(y=c(0,-1024),x=c(0,1024)) +
  geom_hline(data=full_classified_mean_pmnposition, aes(yintercept=-(ave_y+5*sd_y)),linetype='dashed', col = 'red') +
  facet_wrap(~ embryoID) +
  # coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)

```

Identify y pos of cells to be filtered

```{r}

full_classified_to_be_filtered <- full_classified_mean_pmnposition %>%
  mutate(pMN_y5sd_filter=ave_y+5*sd_y) %>% 
  select(ImageNumber,embryoID,pMN_y5sd_filter) %>% unique()
  

```


Re-classify after positional filter

```{r fig.width=12, fig.height=12}

classified_filtered <- full_classified %>%
  left_join(full_classified_to_be_filtered, by = c("ImageNumber","embryoID")) %>% 
  group_by(ImageNumber,embryoID) %>%
  filter(Location_Center_Y < pMN_y5sd_filter)
 
 
ggplot(classified_filtered, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  geom_point(size=0.1, aes(color=celltype))+
  facet_wrap( ~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

ggplot(classified_filtered, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  geom_point(size=0.1, aes(color=foxa2_positive))+
  facet_wrap( ~ embryoID) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  theme_bw()

```


## the plot

```{r}

# classified_filtered

classified_filtered_celltype_counts <- classified_filtered %>%
  mutate(celltype = factor(celltype, levels = c("DP","p3","pMN","Sox2_notp3","othercell"))) %>%
  # mutate(foxa2_positive = factor(foxa2_positive, levels = c("Foxa2","notFoxa2"))) %>%
  group_by(embryoID,genotype) %>%
  count(celltype,.drop=FALSE)

classified_foxa2_counts <- classified_filtered %>%
  # mutate(celltype = factor(celltype, levels = c("DP","p3","pMN","Sox2_notp3","othercell"))) %>%
  mutate(foxa2_positive = factor(foxa2_positive, levels = c("Foxa2","notFoxa2"))) %>%
  group_by(embryoID) %>%
  count(foxa2_positive,.drop=FALSE)

classified_foxa2_counts_wide <- classified_foxa2_counts %>%
  pivot_wider(names_from = "foxa2_positive", values_from = "n")

classified_filtered_celltype_and_FOXA2 <- classified_filtered_celltype_counts %>%
  pivot_wider(names_from = "celltype", values_from = "n") %>%
  left_join(classified_foxa2_counts_wide, by="embryoID") %>%
  separate(embryoID, into = c("block","pos","emb"), sep = "_",remove = FALSE) %>%
  mutate(embryorealID = paste(block,emb, sep="_"))
  
```

No clear trend with just cell numbers 

```{r fig.width=6, fig.height=2.5}


ggplot(classified_filtered_celltype_and_FOXA2, aes(x=Foxa2, y=DP)) +
  geom_point(aes(fill=embryorealID), color="black", shape=21) +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2, aes(x=Foxa2, y=DP)) +
  geom_point(aes(fill=genotype), color="black", shape=21) +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)

```


Is it just bigger embryos have more cells? 

```{r fig.width=6, fig.height=2.5}


ggplot(classified_filtered_celltype_and_FOXA2, aes(x=Foxa2, y=p3)) +
  geom_point(aes(fill=embryorealID), color="black", shape=21) +
  ylab("Number of p3 cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2, aes(x=Foxa2, y=p3)) +
  geom_point(aes(fill=genotype), color="black", shape=21) +
  ylab("Number of p3 cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)

```


Proportions instead?

The amount of total cells considered should be DP+p3+pMN+Sox2_notp3

```{r}

classified_filtered_celltype_and_FOXA2_prop <- classified_filtered_celltype_and_FOXA2 %>%
  mutate(NTcells=sum(DP,p3,pMN,Sox2_notp3)) %>%
  mutate(prop_DP = DP/NTcells) %>%
  mutate(prop_FOXA2 = Foxa2/NTcells)


```

```{r fig.width=6.5, fig.height=2.5}


ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=prop_FOXA2, y=prop_DP)) +
  geom_point(aes(fill=embryorealID), color="black", shape=21) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=prop_FOXA2, y=prop_DP)) +
  geom_point(aes(fill=genotype), color="black", shape=21) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)



```



```{r fig.width=6.5, fig.height=2.5}


ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=prop_FOXA2, y=prop_DP)) +
  geom_point(aes(fill=NTcells), color="black", shape=21) +
  scale_fill_viridis_c(direction=-1) +
  #scale_fill_gradient2(low = "grey", mid = "white", high = "brown") +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)


ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  geom_point(aes(fill=NTcells), color="black", shape=21) +
  scale_fill_viridis_c(direction=-1) +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap( ~pos) +
  theme_bw(base_size = 12)


```


```{r }

classified_filtered_celltype_and_FOXA2_prop_ave <- classified_filtered_celltype_and_FOXA2_prop %>%
  group_by(embryorealID,genotype) %>%
  summarise(mean_propDP=mean(prop_DP),
         sd_propDP=sd(prop_DP),
         sd_propFOXA2=sd(prop_FOXA2),
         mean_propFOXA2=mean(prop_FOXA2),
         mean_NTcell=mean(NTcells))


ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=embryorealID, y=mean_propDP)) +
  geom_col(aes(fill=embryorealID)) +
  geom_point(data = classified_filtered_celltype_and_FOXA2_prop, aes(x=embryorealID, y=prop_DP, fill=embryorealID,shape=pos), color="black") +
  theme_bw() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=embryorealID, y=mean_propFOXA2)) +
  geom_col(aes(fill=embryorealID)) +
  geom_point(data = classified_filtered_celltype_and_FOXA2_prop, aes(x=embryorealID, y=prop_FOXA2, fill=embryorealID,shape=pos), color="black") +
  theme_bw() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=embryorealID, y=mean_propFOXA2)) +
  geom_col(aes(fill=genotype)) +
  geom_point(data = classified_filtered_celltype_and_FOXA2_prop, aes(x=embryorealID, y=prop_FOXA2, fill=genotype,shape=pos), color="black") +
  theme_bw() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=embryorealID, y=mean_NTcell)) +
  geom_col(aes(fill=embryorealID)) +
  geom_point(data = classified_filtered_celltype_and_FOXA2_prop, aes(x=embryorealID, y=NTcells, fill=embryorealID,shape=pos), color="black") +
  theme_bw() +theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r }


ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=embryorealID), color="black", shape=21) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=mean_NTcell), color="black", shape=21) +
  scale_fill_viridis_c(direction=-1) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=genotype), color="black", shape=21) +
  # scale_fill_viridis_c(direction=-1) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  theme_bw(base_size = 12)

```

```{r fig.height=3, fig.width=10}

# 4 bins of equal length

classified_filtered_celltype_and_FOXA2_prop_ave$NTdiscrete <- cut_interval(classified_filtered_celltype_and_FOXA2_prop_ave$mean_NTcell, 4)

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=embryorealID), color="black", shape=21) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=mean_NTcell), color="black", shape=21) +
  scale_fill_viridis_c(direction=-1) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop_ave, aes(x=mean_propFOXA2, y=mean_propDP)) +
  geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=genotype), color="black", shape=21) +
  # scale_fill_viridis_c(direction=-1) +
  ylab("Proportion of DP+ cells per section") +
  xlab("Proportion of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)


```

```{r fig.height=4, fig.width=12}
# classified_filtered_celltype_and_FOXA2_prop

# 4 bins of equal length

classified_filtered_celltype_and_FOXA2_prop$NTdiscrete <- cut_interval(classified_filtered_celltype_and_FOXA2_prop$NTcells, 5)
classified_filtered_celltype_and_FOXA2_prop$embletter <- classified_filtered_celltype_and_FOXA2_prop$emb %>% 
                                              str_replace_all("[1-9]", "")


ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  # geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  # geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=embryorealID), color="black", shape=21, size=3) +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  # geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  # geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=embletter, shape=block),size=3) +
  scale_shape_manual(values = shapes5_manual) +
  scale_fill_brewer(palette = "Set2") +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  # geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  # geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(color=embletter, shape=block), size=3) +
  scale_shape_manual(values = shapes5_manual) +
  scale_color_brewer(palette = "Set2") +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)


ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  # geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  # geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=block, color=block, shape=pos), size=3) +
  #scale_shape_manual(values = shapes5_manual) +
  scale_fill_brewer(palette = "Set2") +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)

ggplot(classified_filtered_celltype_and_FOXA2_prop, aes(x=Foxa2, y=DP)) +
  # geom_linerange(aes(xmin = mean_propFOXA2-sd_propFOXA2,xmax = mean_propFOXA2+sd_propFOXA2), color="#cecece") + 
  # geom_linerange(aes(ymin = mean_propDP+sd_propDP,ymax = mean_propDP-sd_propDP), color="#cecece") + 
  geom_point(aes(fill=genotype, shape=block),size=3) +
  scale_shape_manual(values = shapes5_manual) +
  #scale_fill_brewer(palette = "Set2") +
  ylab("Number of DP+ cells per section") +
  xlab("Number of FOXA2+ cells per section") +
  facet_wrap(~ NTdiscrete, nrow = 1) +
  theme_bw(base_size = 12)


```

## Attempts at plotting Foxa2 levels

```{r fig.width=14, fig.height=12}

ggplot(segment_cp_geno, aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_FOXA2_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered, aes(x=-Location_Center_Y,y=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_SOX2_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered, aes(x=-Location_Center_Y,y=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_NKX22_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "darkgreen") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

```
```{r}
ggplot(classified_filtered %>% filter(embryoID=="8.1fA_Pos2_A1"), aes(y=-Location_Center_Y,x=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=Intensity_MeanIntensity_NKX22_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "darkgreen") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered %>% filter(embryoID=="8.1fA_Pos2_A1"), aes(y=-Location_Center_Y,x=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=celltype))+
  #scale_color_gradient2(low = "grey", mid = "white", high = "grey10") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered %>% filter(embryoID=="8.1fA_Pos2_A1") %>%
         filter(Location_Center_X>500), aes(y=-Location_Center_Y,x=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=celltype))+
  #scale_color_gradient2(low = "grey", mid = "white", high = "grey10") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered %>% filter(embryoID=="8.1fA_Pos2_A1") %>%
         filter(Location_Center_X<500), aes(y=-Location_Center_Y,x=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=0.1, aes(color=celltype))+
  #scale_color_gradient2(low = "grey", mid = "white", high = "grey10") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered %>% filter(embryoID=="8.1fA_Pos2_A1") %>%
         filter(Location_Center_X<500), aes(y=-Location_Center_Y,x=Intensity_MeanIntensity_FOXA2_Image)) +
  #geom_point(size=0.01) +
  geom_point(size=1, aes(color=Intensity_MeanIntensity_FOXA2_Image))+
  scale_color_gradient2(low = "white", mid = "grey90", high = "grey20") +
  facet_wrap( ~ embryoID) +
  ggtitle("X under 500") +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered  %>% filter(embryoID=="8.1fA_Pos2_A1"), aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.5, aes(color=Intensity_MeanIntensity_FOXA2_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "grey20") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

ggplot(classified_filtered  %>% filter(embryoID=="8.1fA_Pos2_A1"), aes(x=Location_Center_X,y=-Location_Center_Y)) +
  #geom_point(size=0.01) +
  geom_point(size=0.5, aes(color=Intensity_MeanIntensity_NKX22_Image))+
  scale_color_gradient2(low = "grey", mid = "white", high = "darkgreen") +
  facet_wrap( ~ embryoID) +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  theme_bw()

```


```{r}
sessionInfo()
```
