---
title: "R_Deseq_PCA_Heatmaps"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RNA analysis April 2024
Re-analysis after nf-core/rnaseq 3.5 

```{r message=FALSE}

rm(list=ls())

library(DESeq2)
library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(tximport)

```




### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="inputs_CaTSRNA/"

outdir="outputs_CaTSRNA_1/"
ifelse(!dir.exists(file.path(workingdir,outdir)), dir.create(file.path(workingdir,outdir)), "Directory exists")



```


## Load data

I am using the output of star_salmon so the import to deseq is a bit more complicated.

```{r }


#salmon counts from pipeline, import
path_files =  list.files(paste0(workingdir,subworkinput)) 
samples = data.frame(run=path_files, stringsAsFactors = FALSE) %>%
  filter(str_detect(run, "^D"))

files <- file.path(paste0(workingdir,subworkinput), samples$run, "quant.sf")
names(files) <- samples$run
all(file.exists(files))

#from pipeline
tx2gene = read_tsv(paste0(paste0(workingdir,subworkinput),"/salmon_tx2gene.tsv"))

txi.salmon <- tximport(files, type = "salmon", tx2gene = tx2gene)
#head(txi.salmon$counts)

```

genes named Geneid


## qc samples - all counts
```{r fig.height=9, fig.width=6}

count.table <- txi.salmon$counts


sums.counts <- count.table %>% as.data.frame() %>%
  gather(sampleid,counts,starts_with("D")) %>%
  group_by(sampleid) %>%
  summarise(totalcounts = sum(counts))


ggplot(sums.counts, aes(y=sampleid, x=totalcounts)) +
  geom_bar(stat="identity")


```



## Colors and shapes
Before more plotting, let's get some metadata organised

```{r }
sorted_gate <- c("pMN","DP","p3","neurons")
sorted_conditions <- c("500","UPSAG","dRA2UPSAG","dRA")

# sorted.DayGate <- c("D5_p1","D5_p2","D5_pM",
#                     "D7_p1","D7_p2","D7_pM",
#                     "D9_p1","D9_p2","D9_pM",
#                     "D11_p1","D11_p2","D11_pM")
# sorted.day <- c("D5","D7","D9","D11")
# sorted.dayNfia <- c("D5_NFIAn_WT","D5_NFIAn_MUT",
#                     "D7_NFIAn_WT","D7_NFIAn_MUT",
#                     "D9_NFIAn_WT","D9_NFIAp_WT","D9_NFIAn_MUT",
#                     "D11_NFIAp_WT","D11_NFIAn_MUT")
# 

# 
# nfiaStatus_order <- c("NFIAn_WT","NFIAp_WT","NFIAn_MUT")

shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
 

color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")

# conditions

colors_conditions <- c("#e67300","#4d9a00","#cdcd00","#0073e6")

```







## Differential analysis

This normalization includes ERCC. I use it later samples don't cluster in any special way just with ERCC correlations. 

```{r}


count.matrix <- count.table

## Make metadata file for DESeq

genecolData_first <- data.frame(Sample_ID = colnames(count.matrix))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))


dds <- DESeqDataSetFromTximport(txi = txi.salmon,
                              colData = genecolData_first,
                              design = Condition ~ Gate)

dds <- DESeq(dds)

vsd <- varianceStabilizingTransformation(dds,blind = FALSE)

# Export normalized tables for plotting elsewhere
dds_counts <- counts(dds, normalized = TRUE)
vsd_data <- assay(vsd)

```

Removing Outliers

#```{r}

vsd_data <- subset(vsd_data_unfiltered, select = -D6_UPSAG_pMN_R2)

```

## Plot PCAs 

```{r }
## to get more than 2 components

## to go it in the top ntop variable genes 
# calculate the variance for each gene
rv <- rowVars(vsd_data)
# select the ntop genes by variance
ntop=500
select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]

#pca <- prcomp(t(assay(object)[select,]))
t_vsd <- t(vsd_data[select,])
vsd_pca <- prcomp(t_vsd, retx=TRUE, scale. = FALSE)

names(vsd_pca)

# How many PC to explain enough variance?
#summary(vsd_pca)

var_explained <-vsd_pca$sdev^2/sum(vsd_pca$sdev^2)
plot(var_explained)

```



```{r }

vsd_pca_plot <- vsd_pca$x %>% 
  as.data.frame %>%
  rownames_to_column("Sample") %>%
  separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Experiment=paste(Condition,Rep,sep="_"),
         Gate=factor(Gate, levels=sorted_gate))
  
ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Day,shape=Gate)) +
  scale_fill_manual(values = colors_greys) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Day)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Condition,shape=Gate)) +
  scale_fill_manual(values = colors_conditions) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=DayGate,shape=Condition)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=DayCondition,shape=Gate)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Rep)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```



## Plot more components

```{r }

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Condition)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Day,shape=Condition)) +
  scale_fill_manual(values = colors_greys) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC2,y=PC3,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)


ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Rep,shape=Gate, label=Condition)) +
  geom_point(size=4, alpha=0.9) +
  geom_text(nudge_x = 5) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)


ggplot(vsd_pca_plot, aes(x=PC4,y=PC2,fill=Condition,shape=Day)) +
  #scale_fill_manual(values = colorIZ) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC4: ",round(var_explained[4]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)


ggplot(vsd_pca_plot, aes(x=PC4,y=PC5,fill=DayCondition,shape=Gate, label=Gate)) +
  geom_point(size=4, alpha=0.9) +
  geom_text(nudge_x = 5) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC4: ",round(var_explained[4]*100,1),"%"),
       y=paste0("PC5: ",round(var_explained[5]*100,1),"%")) +
  theme_bw(base_size=16)


```



# Correlation Heatmaps - based on 500 top genes 
```{r fig.width=11,fig.height=11, eval=FALSE, include=FALSE}

cor_vsd <-cor( vsd_data[select,] )

names(color_gates) <- sorted_gate
names(colors_conditions) <- sorted_conditions

phen_data <- genecolData_first %>%
  select(c("Sample_ID","Day","Condition","Gate","Rep")) %>%
  remove_rownames() %>%
  column_to_rownames("Sample_ID")
ann_color_JD <- list(Gate=color_gates, Condition=colors_conditions,
    Day = c(D5="#f6f6f6",D6="#808080",D7="#333333"),
  Rep = c(R1="#ebeb77", R2="#77b1eb"))


# Build the annotation for the complex heatmap
heatmap_ann_row <- rowAnnotation(df=phen_data, col=ann_color_JD)
heatmap_ann <- HeatmapAnnotation(df=phen_data, col=ann_color_JD, show_legend = FALSE)



# Annotated heatmap with selected colors
hm_colors = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100)


Heatmap(cor_vsd, name="Correlation", col=hm_colors,
        cluster_columns = TRUE, cluster_rows = TRUE,
        #show_column_dend = FALSE,
        #row_dend_side = "right", column_dend_side = "bottom",
        # cluster methods for rows and columns
        clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
        clustering_method_columns = 'ward.D2',
        clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
        clustering_method_rows = 'ward.D2',
        column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"),
        row_names_gp = gpar(fontsize = 10),column_names_gp = gpar(fontsize = 10),
        left_annotation = heatmap_ann_row, top_annotation = heatmap_ann,
        column_title = "Pearson Correlation - Top 500 genes")



```


## Export files 

```{r }
# Export normalized tables for plotting elsewhere
dds_counts <- counts(dds, normalized = TRUE)
vsd_data <- assay(vsd)

write.table(dds_counts, file = paste0(workingdir,outdir,"featurecounts.normCounts.txt"), quote = FALSE, row.names = TRUE)
write.csv(vsd_data, file = paste0(workingdir,outdir,"featurecounts.vsd.csv"), quote = FALSE)

```



## Genes to look at

```{r}

dds_counts_plot <- dds_counts %>% 
  as.data.frame() %>%
  rownames_to_column("geneid") %>%
  gather(sampleid, counts_norm, starts_with("D")) %>%
  separate(sampleid,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Gate=factor(Gate, levels=sorted_gate),
         Condition=factor(Condition, levels=sorted_conditions))
```


```{r fig.height=9, fig.width=9}

geneOI <- c("Sox2","Dbx1","Irx3","Pax6","Nkx6-1","Olig2","Nkx2-2","Foxa2","Shh","Arx","Tubb3","Sim1","Mnx1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```
```{r fig.height=4, fig.width=5}

geneOI_2 <- c("Nkx2-9")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI_2) %>% mutate(geneid=factor(geneid, levels=geneOI_2)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free_y") +
  theme_bw()
```

# RA Response Genes

```{r fig.height=10, fig.width=8}

geneOI_RAR <- c("Rara","Rarb","Rarg","Cyp26b1","Cyp26a1","Cyp26c1","Crabp2","Crabp1","Aldh1a2","Aldh1a1","Aldh1a3","Capn6","Nrarp","Pard6g","Mrpl41","Smu1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI_RAR) %>% mutate(geneid=factor(geneid, levels=geneOI_RAR)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free_y") +
  theme_bw()
```

# Analysis without neurons

```{r}

gates=c("_pMN_","_p3_","_DP_")

sub_counts <- count.table %>% 
  as.data.frame() %>%
  dplyr::select(contains(gates))

## Make metadata file for DESeq

genecolData_first <- data.frame(Sample_ID = colnames(sub_counts))

genecolData_first <- genecolData_first %>% 
  separate(Sample_ID,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) 

genecolData_first <- as.data.frame(unclass(genecolData_first))


dds <- DESeqDataSetFromMatrix(countData =  round(sub_counts),
                                  colData = genecolData_first,
                                  design = ~ Gate)
dds <- DESeq(dds)

vsd <- varianceStabilizingTransformation(dds,blind = FALSE)

# Export normalized tables for plotting elsewhere
dds_counts <- counts(dds, normalized = TRUE)
vsd_data <- assay(vsd)

```

## Plot PCAs 

```{r }
## to get more than 2 components

## to go it in the top ntop variable genes 
# calculate the variance for each gene
rv <- rowVars(vsd_data)
# select the ntop genes by variance
ntop=500
select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]

#pca <- prcomp(t(assay(object)[select,]))
t_vsd <- t(vsd_data[select,])
vsd_pca <- prcomp(t_vsd, retx=TRUE, scale. = FALSE)

names(vsd_pca)

# How many PC to explain enough variance?
#summary(vsd_pca)

var_explained <-vsd_pca$sdev^2/sum(vsd_pca$sdev^2)
plot(var_explained)

```

```{r }

vsd_pca_plot <- vsd_pca$x %>% 
  as.data.frame %>%
  rownames_to_column("Sample") %>%
  separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Experiment=paste(Condition,Rep,sep="_"),
         Gate=factor(Gate, levels=sorted_gate))
  
ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Day,shape=Gate)) +
  scale_fill_manual(values = colors_greys) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Day)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Condition,shape=Gate)) +
  scale_fill_manual(values = colors_conditions) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=DayGate,shape=Condition)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=DayCondition,shape=Gate)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Rep)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

```
## More components

```{r }

ggplot(vsd_pca_plot, aes(x=PC1,y=PC2,fill=Gate,shape=Condition)) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Day,shape=Condition)) +
  scale_fill_manual(values = colors_greys) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC2,y=PC3,fill=Gate,shape=Condition)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)
ggplot(vsd_pca_plot, aes(x=PC2,y=PC3,fill=Gate,shape=Condition,label=Sample)) +
  scale_fill_manual(values = color_gates) +
  geom_point(size=4, alpha=0.9) +
  geom_text(nudge_x = 10) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC2: ",round(var_explained[2]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)

ggplot(vsd_pca_plot, aes(x=PC1,y=PC3,fill=Rep,shape=Gate, label=Condition)) +
  geom_point(size=4, alpha=0.9) +
  geom_text(nudge_x = 5) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC1: ",round(var_explained[1]*100,1),"%"),
       y=paste0("PC3: ",round(var_explained[3]*100,1),"%")) +
  theme_bw(base_size=16)


ggplot(vsd_pca_plot, aes(x=PC4,y=PC2,fill=Condition,shape=Day)) +
  #scale_fill_manual(values = colorIZ) +
  geom_point(size=4, alpha=0.9) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes5_manual) +
  labs(x=paste0("PC4: ",round(var_explained[4]*100,1),"%"),
       y=paste0("PC2: ",round(var_explained[2]*100,1),"%")) +
  theme_bw(base_size=16)


ggplot(vsd_pca_plot, aes(x=PC4,y=PC5,fill=DayCondition,shape=Gate, label=Gate)) +
  geom_point(size=4, alpha=0.9) +
  geom_text(nudge_x = 5) +
  guides(fill = guide_legend(override.aes=list(shape=21))) +
  scale_shape_manual(values = shapes4_fill_manual) +
  labs(x=paste0("PC4: ",round(var_explained[4]*100,1),"%"),
       y=paste0("PC5: ",round(var_explained[5]*100,1),"%")) +
  theme_bw(base_size=16)


```


```{r}
sessionInfo()
```


