---
title: "RNA_3_diffexpression"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# RNA analysis
Differential expression between gates, conditions and timepoints

```{r message=FALSE}

rm(list=ls())

library(DESeq2)
library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(tximport)
library(ggrepel)

```


### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="inputs_RNApilot/"
outdir="outputs_RNApilot_3_diffexpression/"

ifelse(!dir.exists(file.path(workingdir,outdir)), dir.create(file.path(workingdir,outdir)), "Directory exists")

suboutdir1="output_between_gates/"
ifelse(!dir.exists(file.path(workingdir,outdir,suboutdir1)), dir.create(file.path(workingdir,outdir,suboutdir1)), "Directory exists")
suboutdir2="output_between_timepoints/"
ifelse(!dir.exists(file.path(workingdir,outdir,suboutdir2)), dir.create(file.path(workingdir,outdir,suboutdir2)), "Directory exists")
suboutdir3="output_between_conditions/"
ifelse(!dir.exists(file.path(workingdir,outdir,suboutdir3)), dir.create(file.path(workingdir,outdir,suboutdir3)), "Directory exists")


```



## Load data

For RNA analysis, we are using the output of star_salmon so the import to deseq is a bit more complicated.

```{r }

#salmon counts from pipeline, import
path_files =  list.files(paste0(workingdir,subworkinput)) 
samples = data.frame(run=path_files, stringsAsFactors = FALSE) %>%
  filter(str_detect(run, "^D"))

files <- file.path(paste0(workingdir,subworkinput), samples$run, "quant.sf")
names(files) <- samples$run
all(file.exists(files))

#from pipeline
tx2gene = read_tsv(paste0(paste0(workingdir,subworkinput),"/salmon_tx2gene.tsv"))

txi.salmon <- tximport(files, type = "salmon", tx2gene = tx2gene)


```



## Colors and shapes

```{r }
sorted_gate <- c("pMN","DP","p3","neur")
sorted_conditions <- c("500","UPSAG","dRA2UPSAG","dRA")


shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
 

color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")

# conditions

colors_conditions <- c("#e67300","#4d9a00","#cdcd00","#0073e6")
```

## Differential analysis between domains for a given timepoint and condition

The lapply is not running yet because I need to filter the comparisons with enough samples. 

Targeted diff analysis in subsets of samples:
- for each timepoint: diff expression between domains

- make a extra lapply to go through conditions

This DESeq2 analysis is done by subsetting samples. I subset the `txi.salmon$counts` table and then use ` DESeqDataSetFromMatrix` with `round(sub_counts)` to perform the differential analysis. 

```{r message = FALSE, warning=FALSE}

count_matrix <- txi.salmon$counts %>%
  as.data.frame()


#subset 
treatment=c("_500_","_UPSAG_","_dRA_","_dRA2UPSAG_")

timepoint=c("D5_","D6_","D7_")

#matrix
allgates=matrix(c("_pMN_","_p3_","_pMN_","_DP_","_p3_","_DP_"),
                 nrow=2,
                 ncol=3)

comparisons <- allgates

# count_matrix <- count_table %>%
#   column_to_rownames("Geneid")


# x=3
# y=2
# i=3

## many comparisons do not have 2 reps and 2 conditions

PairWiseDEseq <-  lapply(c(1:length(treatment)),function (i) {
  lapply(c(1:length(timepoint)),function (x) {
    lapply(c(1:ncol(allgates)), function (y) {
        regime <- treatment[i]
        timepoints <- timepoint[x]
        celltypes <- allgates[,y]
        sub_counts <- count_matrix %>%
          dplyr::select(contains(regime)  & contains(celltypes) & contains(timepoints))
        
      ## if_else statement because many comparisons do not have enough samples
        
        if(ncol(sub_counts)==4){
      
          ## Make metadata file for DESeq
          genecolData_sub <- data.frame(Sample = colnames(sub_counts))
          genecolData_sub <- genecolData_sub %>% 
            separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
            mutate(DayGate=factor(paste(Day,Gate,sep="_")),
                  DayCondition=paste(Day,Condition),
                  Experiment=paste(Condition,Rep,sep="_"),
                  Gate=factor(Gate, levels=sorted_gate))
          genecolData_sub <- as.data.frame(unclass(genecolData_sub))
          
          
          dds_sub <- DESeqDataSetFromMatrix(countData =  round(sub_counts),
                                  colData = genecolData_sub,
                                  design = ~ Gate)
              
          dds_sub <- DESeq(dds_sub)
          
          vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
          
          # Export normalized tables for plotting elsewhere
          dds_sub_counts <- counts(dds_sub, normalized = TRUE)
          vsd_sub_data <- assay(vsd_sub)
          
          results_sub <- results(dds_sub)
    
          # color significant
          results_sub_plot1 <- results_sub %>%
            as.data.frame() %>%
            rownames_to_column("geneid") %>%
            mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                       padj < 0.01 & padj > 0.001 ~ "under001",
                                       padj < 0.001 & padj >0 ~ "under0001",
                                       TRUE ~ "over01"))
          
          #threshold <- 8
          plot_ma <- ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
              geom_point(size=1) +
              # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
              #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
              # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
              #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
              #ylim(-threshold,threshold) +
              scale_x_log10() +
              scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
              geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                      nudge_y = 10,
                      #nudge_x=-10,
                      #force_pull   = 10,
                      force         = 65,
                      max.overlaps = Inf,
                      box.padding = 0.5,
                      segment.color = "grey50",
                      direction     = "both") +
              geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                      nudge_y = -20,
                      #nudge_x=-10,
                      #force_pull   = 10,
                      force         = 65,
                      max.overlaps = Inf,
                      box.padding = 0.5,
                      segment.color = "grey50",
                      direction     = "both") +
              ylab(paste0("log2 Fold Change ", resultsNames(dds_sub)[2] ," RNA")) +
              ggtitle(paste0("Results ",regime,timepoints,resultsNames(dds_sub)[2])) +
              theme_bw() 
          
          # print plot
          ggsave(paste0(workingdir,outdir,suboutdir1,"MAplot_",regime,timepoints,resultsNames(dds_sub)[2],".pdf"), plot=plot_ma,
           width=5, height=3, units="in", useDingbats=FALSE)
          
    
          ## Export files
          
          write.table(dds_sub_counts,
          file = paste0(workingdir,outdir,suboutdir1,"CountsNormalized_",regime,timepoints,resultsNames(dds_sub)[2],".txt"),
              quote = FALSE, row.names = TRUE)
          write.csv(vsd_sub_data,
              paste0(workingdir,outdir,suboutdir1,"VSData_",regime,timepoints,resultsNames(dds_sub)[2],".csv"),
              quote = FALSE)
          write.table(results_sub,
              file = paste0(workingdir,outdir,suboutdir1,"Results_DESeq_",regime,timepoints,resultsNames(dds_sub)[2],".txt"),
              quote = FALSE, row.names = TRUE)
    
          results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
          #results_return$Comparison <- paste0("Comp_",timepoints,celltypes[1],celltypes[2])
          results_return$Comparison <- paste0("Comp_",regime,timepoints,resultsNames(dds_sub)[2])
          results_return
        } 
  
    })
  })
})

```

## Differential analysis between timepoints for each domain and condition 

For each gate:
- differential analysis between timepoints for the same condition: To do, extra lapply for conditions 



```{r message = FALSE, warning=FALSE}

#subset 
treatment=c("_500_","_UPSAG_","_dRA_","_dRA2UPSAG_")

allgates=c("_pMN_","_p3_","_DP_")
#matrix
timepoints=matrix(c("D5_","D6_","D5_","D7_","D6_","D7_"),
                 nrow=2,
                 ncol=3)

comparisons <- timepoints

# i=1
# x=2
# y=3

PairWiseDEseq <- lapply(c(1:length(treatment)),function (i) {
  lapply(c(1:length(allgates)),function (x) {
    lapply(c(1:ncol(timepoints)), function (y) {
        regime <- treatment[i]
        celltypes <- allgates[x]
        days <- timepoints[,y]
        sub_counts <- count_matrix %>%
          dplyr::select(contains(regime)  & contains(celltypes) & contains(days))
        
         ## if_else statement because many comparisons do not have enough samples
        
        if(ncol(sub_counts)==4){
        
        ## Make metadata file for DESeq
        genecolData_sub <- data.frame(Sample = colnames(sub_counts))
        genecolData_sub <- genecolData_sub %>% 
          separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
          mutate(DayGate=factor(paste(Day,Gate,sep="_")),
                DayCondition=paste(Day,Condition),
                Experiment=paste(Condition,Rep,sep="_"),
                Gate=factor(Gate, levels=sorted_gate))
        genecolData_sub <- as.data.frame(unclass(genecolData_sub))
           
        dds_sub <- DESeqDataSetFromMatrix(countData = round(sub_counts),
                                      colData = genecolData_sub,
                                      design = ~ Day)
        
        dds_sub <- DESeq(dds_sub)
        
        vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
        
        # Export normalized tables for plotting elsewhere
        dds_sub_counts <- counts(dds_sub, normalized = TRUE)
        vsd_sub_data <- assay(vsd_sub)
        
        results_sub <- results(dds_sub)
        
        ## plot MA custom plot 
        
        #plotMA(results_sub,ylim=c(-8,8))
        
        # color significant
        results_sub_plot1 <- results_sub %>%
          as.data.frame() %>%
          rownames_to_column("geneid") %>%
          mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                     padj < 0.01 & padj > 0.001 ~ "under001",
                                     padj < 0.001 & padj >0 ~ "under0001",
                                     TRUE ~ "over01"))
        
        #threshold <- 8
         plot_ma <- ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
            geom_point(size=1) +
            # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
            #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
            # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
            #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
            #ylim(-threshold,threshold) +
            scale_x_log10() +
            scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
            geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                    nudge_y = 10,
                    #nudge_x=-10,
                    #force_pull   = 10,
                    force         = 65,
                    max.overlaps = Inf,
                    box.padding = 0.5,
                    segment.color = "grey50",
                    direction     = "both") +
            geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                    nudge_y = -20,
                    #nudge_x=-10,
                    #force_pull   = 10,
                    force         = 65,
                    max.overlaps = Inf,
                    box.padding = 0.5,
                    segment.color = "grey50",
                    direction     = "both") +
            ylab(paste0("log2 Fold Change ",resultsNames(dds_sub)[2]," RNA")) +
            ggtitle(paste0("Results ",regime,celltypes,resultsNames(dds_sub)[2] )) +
            theme_bw() 
        
         # print plot
          ggsave(paste0(workingdir,outdir,suboutdir2,"MAplot_",regime,celltypes,resultsNames(dds_sub)[2],".pdf"), plot=plot_ma,
           width=5, height=3, units="in", useDingbats=FALSE)
        
        ## Export files
        
        write.table(dds_sub_counts,
        file = paste0(workingdir,outdir,suboutdir2,"CountsNormalized_",regime,celltypes,resultsNames(dds_sub)[2],".txt"),
            quote = FALSE, row.names = TRUE)
        write.csv(vsd_sub_data,
            paste0(workingdir,outdir,suboutdir2,"VSData_",regime,celltypes,resultsNames(dds_sub)[2],".csv"),
            quote = FALSE)
        write.table(results_sub,
            file = paste0(workingdir,outdir,suboutdir2,"Results_DESeq_",regime,celltypes,resultsNames(dds_sub)[2],".txt"),
            quote = FALSE, row.names = TRUE)

        results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
        results_return$Comparison <- paste0("Comp_",regime,celltypes,resultsNames(dds_sub)[2])
        results_return
        }
    })
  }) 
}) 

```

## Differential analysis between conditions for each domain and timepoint 

For a given timepoint
- differential analysis between conditions each cell type


```{r message = FALSE, warning=FALSE}

#subset
timepoint=c("D5_","D6_","D7_")
allgates=c("_pMN_","_p3_","_DP_","_neur_")
#matrix
comparisons=matrix(c("_500_","_UPSAG_","_UPSAG_","_dRA2UPSAG_","_dRA2UPSAG_","_dRA_",
                    "_500_","_dRA2UPSAG_","_500_","_dRA_","_UPSAG_","_dRA_"),
                 nrow=2,
                 ncol=6)


# i=3
# x=2
# y=5

PairWiseDEseq <-  lapply(c(1:length(timepoint)),function (i) {
  lapply(c(1:length(allgates)),function (x) {
    lapply(c(1:ncol(comparisons)), function (y) {
        timepoints <- timepoint[i]
        celltypes <- allgates[x]
        conditions <- comparisons[,y]
        sub_counts <- count_matrix %>%
          dplyr::select(contains(timepoints)  & contains(celltypes) & contains(conditions))
        
         ## if_else statement because many comparisons do not have enough samples
        if(ncol(sub_counts)==4){
          
        ## Make metadata file for DESeq
        genecolData_sub <- data.frame(Sample = colnames(sub_counts))
        genecolData_sub <- genecolData_sub %>% 
          separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
          mutate(DayGate=factor(paste(Day,Gate,sep="_")),
                DayCondition=paste(Day,Condition),
                Experiment=paste(Condition,Rep,sep="_"),
                Gate=factor(Gate, levels=sorted_gate))
        genecolData_sub <- as.data.frame(unclass(genecolData_sub))
           
        dds_sub <- DESeqDataSetFromMatrix(countData = round(sub_counts),
                                      colData = genecolData_sub,
                                      design = ~ Condition)
        
        dds_sub <- DESeq(dds_sub)
        
        vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
        
        # Export normalized tables for plotting elsewhere
        dds_sub_counts <- counts(dds_sub, normalized = TRUE)
        vsd_sub_data <- assay(vsd_sub)
        
        results_sub <- results(dds_sub)
        
        ## plot MA custom plot 
        
        #plotMA(results_sub,ylim=c(-8,8))
        
        # color significant
        results_sub_plot1 <- results_sub %>%
          as.data.frame() %>%
          rownames_to_column("geneid") %>%
          mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                     padj < 0.01 & padj > 0.001 ~ "under001",
                                     padj < 0.001 & padj >0 ~ "under0001",
                                     TRUE ~ "over01"))
        
        #threshold <- 8
        plot_ma <-ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
            geom_point(size=1) +
            # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
            #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
            # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
            #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
            #ylim(-threshold,threshold) +
            scale_x_log10() +
            scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
            geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                    nudge_y = 20,
                    #nudge_x=-10,
                    #force_pull   = 10,
                    force         = 65,
                    max.overlaps = Inf,
                    box.padding = 0.5,
                    segment.color = "grey50",
                    direction     = "both") +
            geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                    nudge_y = -20,
                    #nudge_x=-10,
                    #force_pull   = 10,
                    force         = 65,
                    max.overlaps = Inf,
                    box.padding = 0.5,
                    segment.color = "grey50",
                    direction     = "both") +
            ylab(paste0("log2 Fold Change ",resultsNames(dds_sub)[2]," RNA")) +
            ggtitle(paste0("Results ",timepoints,celltypes,resultsNames(dds_sub)[2])) +
            theme_bw() 
        
        # print plot
          ggsave(paste0(workingdir,outdir,suboutdir3,"MAplot_",timepoints,celltypes,resultsNames(dds_sub)[2],".pdf"), plot=plot_ma,
           width=5, height=3, units="in", useDingbats=FALSE)
        
        
        ## Export files
        
        write.table(dds_sub_counts,
        file = paste0(workingdir,outdir,suboutdir3,"CountsNormalized_",timepoints,celltypes,resultsNames(dds_sub)[2],".txt"),
            quote = FALSE, row.names = TRUE)
        write.csv(vsd_sub_data,
            paste0(workingdir,outdir,suboutdir3,"VSData_",timepoints,celltypes,resultsNames(dds_sub)[2],".csv"),
            quote = FALSE)
        write.table(results_sub,
            file = paste0(workingdir,outdir,suboutdir3,"Results_DESeq_",timepoints,celltypes,resultsNames(dds_sub)[2],".txt"),
            quote = FALSE, row.names = TRUE)

        results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
        results_return$Comparison <- paste0("Comp_",timepoints,celltypes,resultsNames(dds_sub)[2])
        results_return
        }
    })
  })
})

```



```{r}
sessionInfo()
```