---
title: "RNA_3_diffexpression"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# RNA analysis
Differential expression between gates, conditions and timepoints

```{r message=FALSE}

rm(list=ls())

library(DESeq2)
library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)
library(tximport)
library(ggrepel)

```


### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="inputs_RNApilot/"
outdir="outputs_RNApilot_3_diffexpression/"

ifelse(!dir.exists(file.path(workingdir,outdir)), dir.create(file.path(workingdir,outdir)), "Directory exists")

# suboutdir1="output_Domain_Specific/"
# ifelse(!dir.exists(file.path(workingdir,outdir,suboutdir1)), dir.create(file.path(workingdir,outdir,suboutdir1)), "Directory exists")
# suboutdir2="output_Time_Specific/"
# ifelse(!dir.exists(file.path(workingdir,outdir,suboutdir2)), dir.create(file.path(workingdir,outdir,suboutdir2)), "Directory exists")
# 

```



## Load data

For RNA analysis, we are using the output of star_salmon so the import to deseq is a bit more complicated.

```{r }

#salmon counts from pipeline, import
path_files =  list.files(paste0(workingdir,subworkinput)) 
samples = data.frame(run=path_files, stringsAsFactors = FALSE) %>%
  filter(str_detect(run, "^D"))

files <- file.path(paste0(workingdir,subworkinput), samples$run, "quant.sf")
names(files) <- samples$run
all(file.exists(files))

#from pipeline
tx2gene = read_tsv(paste0(paste0(workingdir,subworkinput),"/salmon_tx2gene.tsv"))

txi.salmon <- tximport(files, type = "salmon", tx2gene = tx2gene)


```



## Colors and shapes

```{r }
sorted_gate <- c("pMN","DP","p3","neur")
sorted_conditions <- c("500","UPSAG","dRA2UPSAG","dRA")


shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
 

color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")

# conditions

colors_conditions <- c("#e67300","#4d9a00","#cdcd00","#0073e6")
```

## Differential analysis between domains for a given timepoint and condition

The lapply is not running yet because I need to filter the comparisons with enough samples. 

Targeted diff analysis in subsets of samples:
- for each timepoint: diff expression between domains

- make a extra lapply to go through conditions

This DESeq2 analysis is done by subsetting samples. I subset the `txi.salmon$counts` table and then use ` DESeqDataSetFromMatrix` with `round(sub_counts)` to perform the differential analysis. 

```{r}

count_matrix <- txi.salmon$counts %>%
  as.data.frame()


#subset 
timepoint=c("D5_","D6_","D7_")
#matrix
allgates=matrix(c("_pMN_","_p3_","_pMN_","_DP_","_p3_","_DP_"),
                 nrow=2,
                 ncol=3)
geno="_500_"

comparisons <- allgates

# count_matrix <- count_table %>%
#   column_to_rownames("Geneid")


x=3
y=2

## many comparisons do not have 2 reps and 2 conditions

# PairWiseDEseq <- lapply(c(1:length(timepoint)),function (x) {
#   lapply(c(1:ncol(allgates)), function (y) {
      timepoints <- timepoint[x]
      celltypes <- allgates[,y]
      sub_counts <- count_matrix %>%
        dplyr::select(contains(geno)  & contains(celltypes) & contains(timepoints))
      
      ## Make metadata file for DESeq
      genecolData_sub <- data.frame(Sample = colnames(sub_counts))
      genecolData_sub <- genecolData_sub %>% 
        separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
        mutate(DayGate=factor(paste(Day,Gate,sep="_")),
              DayCondition=paste(Day,Condition),
              Experiment=paste(Condition,Rep,sep="_"),
              Gate=factor(Gate, levels=sorted_gate))
      genecolData_sub <- as.data.frame(unclass(genecolData_sub))
      
      
      dds_sub <- DESeqDataSetFromMatrix(countData =  round(sub_counts),
                              colData = genecolData_sub,
                              design = ~ Gate)
          
      dds_sub <- DESeq(dds_sub)
      
      vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
      
      # Export normalized tables for plotting elsewhere
      dds_sub_counts <- counts(dds_sub, normalized = TRUE)
      vsd_sub_data <- assay(vsd_sub)
      
      results_sub <- results(dds_sub)

      # color significant
      results_sub_plot1 <- results_sub %>%
        as.data.frame() %>%
        rownames_to_column("geneid") %>%
        mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                   padj < 0.01 & padj > 0.001 ~ "under001",
                                   padj < 0.001 & padj >0 ~ "under0001",
                                   TRUE ~ "over01"))
      
      #threshold <- 8
      ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
          geom_point(size=1) +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
          #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
          #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
          #ylim(-threshold,threshold) +
          scale_x_log10() +
          scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                  nudge_y = 10,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                  nudge_y = -20,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          ylab(paste0("log2 Fold Change ",timepoints , celltypes[2]," vs ",celltypes[1] ," RNA")) +
          ggtitle(paste0("Results ",timepoints , celltypes[2]," vs ",celltypes[1])) +
          theme_bw() 
  

      ## Export files
      
      # write.table(dds_sub_counts,
      # file = paste0(workingdir,outdir,suboutdir1,"CountsNormalized_",timepoints,celltypes[1],celltypes[2],".txt"),
      #     quote = FALSE, row.names = TRUE)
      # write.csv(vsd_sub_data,
      #     paste0(workingdir,outdir,suboutdir1,"VSData_",timepoints,celltypes[1],celltypes[2],".csv"),
      #     quote = FALSE)
      # write.table(results_sub,
      #     file = paste0(workingdir,outdir,suboutdir1,"Results_DESeq_",timepoints,celltypes[1],celltypes[2],".txt"),
      #     quote = FALSE, row.names = TRUE)

      results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
      results_return$Comparison <- paste0("Comp_",timepoints,celltypes[1],celltypes[2])
#       results_return
# 
#   })
# }) 


```

## Differential analysis between timepoints for each domain and condition 

For each gate:
- differential analysis between timepoints for the same condition: To do, extra lapply for conditions 



```{r}

#subset 
allgates=c("_pMN_","_p3_","_DP_")
#matrix
timepoints=matrix(c("D5_","D6_","D5_","D7_","D6_","D7_"),
                 nrow=2,
                 ncol=3)
geno="_500_"

comparisons <- timepoints

x=2
y=3

# PairWiseDEseq <- lapply(c(1:length(allgates)),function (x) {
#   lapply(c(1:ncol(timepoints)), function (y) {
      celltypes <- allgates[x]
      days <- timepoints[,y]
      sub_counts <- count_matrix %>%
        dplyr::select(contains(geno)  & contains(celltypes) & contains(days))
      
      ## Make metadata file for DESeq
      genecolData_sub <- data.frame(Sample = colnames(sub_counts))
      genecolData_sub <- genecolData_sub %>% 
        separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
        mutate(DayGate=factor(paste(Day,Gate,sep="_")),
              DayCondition=paste(Day,Condition),
              Experiment=paste(Condition,Rep,sep="_"),
              Gate=factor(Gate, levels=sorted_gate))
      genecolData_sub <- as.data.frame(unclass(genecolData_sub))
         
      dds_sub <- DESeqDataSetFromMatrix(countData = round(sub_counts),
                                    colData = genecolData_sub,
                                    design = ~ Day)
      
      dds_sub <- DESeq(dds_sub)
      
      vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
      
      # Export normalized tables for plotting elsewhere
      dds_sub_counts <- counts(dds_sub, normalized = TRUE)
      vsd_sub_data <- assay(vsd_sub)
      
      results_sub <- results(dds_sub)
      
      ## plot MA custom plot 
      
      #plotMA(results_sub,ylim=c(-8,8))
      
      # color significant
      results_sub_plot1 <- results_sub %>%
        as.data.frame() %>%
        rownames_to_column("geneid") %>%
        mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                   padj < 0.01 & padj > 0.001 ~ "under001",
                                   padj < 0.001 & padj >0 ~ "under0001",
                                   TRUE ~ "over01"))
      
      #threshold <- 8
      ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
          geom_point(size=1) +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
          #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
          #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
          #ylim(-threshold,threshold) +
          scale_x_log10() +
          scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                  nudge_y = 10,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                  nudge_y = -20,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          ylab(paste0("log2 Fold Change ",celltypes , days[2]," vs ",days[1] ," RNA")) +
          ggtitle(paste0("Results ",celltypes , days[2]," vs ",days[1] )) +
          theme_bw() 
      
      
      ## Export files
      
      # write.table(dds_sub_counts,
      # file = paste0(workingdir,outdir,suboutdir2,"CountsNormalized_",celltypes,days[1],days[2],".txt"),
      #     quote = FALSE, row.names = TRUE)
      # write.csv(vsd_sub_data,
      #     paste0(workingdir,outdir,suboutdir2,"VSData_",celltypes,days[1],days[2],".csv"),
      #     quote = FALSE)
      # write.table(results_sub,
      #     file = paste0(workingdir,outdir,suboutdir2,"Results_DESeq_",celltypes,days[1],days[2],".txt"),
      #     quote = FALSE, row.names = TRUE)

      results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
      results_return$Comparison <- paste0("Comp_",celltypes,days[1],days[2])
      #results_return

#   })
# }) 


```

## Differential analysis between conditions for each domain and timepoint 

Ongoing

For a given timepoint
- differential analysis between conditions each cell type


```{r}

#subset 
allgates=c("_pMN_","_p3_","_DP_")
#matrix
comparisons=matrix(c("_dRA_","_500_","_dRA_","_dRA2UPSAG_","_500_","_dRA2UPSAG_"),
                 nrow=2,
                 ncol=3)
timepoint="D7_"



x=2
y=1

# PairWiseDEseq <- lapply(c(1:length(allgates)),function (x) {
#   lapply(c(1:ncol(timepoints)), function (y) {
      celltypes <- allgates[x]
      conditions <- comparisons[,y]
      sub_counts <- count_matrix %>%
        dplyr::select(contains(timepoint)  & contains(celltypes) & contains(conditions))
      
      ## Make metadata file for DESeq
      genecolData_sub <- data.frame(Sample = colnames(sub_counts))
      genecolData_sub <- genecolData_sub %>% 
        separate(Sample,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
        mutate(DayGate=factor(paste(Day,Gate,sep="_")),
              DayCondition=paste(Day,Condition),
              Experiment=paste(Condition,Rep,sep="_"),
              Gate=factor(Gate, levels=sorted_gate))
      genecolData_sub <- as.data.frame(unclass(genecolData_sub))
         
      dds_sub <- DESeqDataSetFromMatrix(countData = round(sub_counts),
                                    colData = genecolData_sub,
                                    design = ~ Condition)
      
      dds_sub <- DESeq(dds_sub)
      
      vsd_sub <- varianceStabilizingTransformation(dds_sub,blind = FALSE)
      
      # Export normalized tables for plotting elsewhere
      dds_sub_counts <- counts(dds_sub, normalized = TRUE)
      vsd_sub_data <- assay(vsd_sub)
      
      results_sub <- results(dds_sub)
      
      ## plot MA custom plot 
      
      #plotMA(results_sub,ylim=c(-8,8))
      
      # color significant
      results_sub_plot1 <- results_sub %>%
        as.data.frame() %>%
        rownames_to_column("geneid") %>%
        mutate(color_sig=case_when(padj < 0.1 & padj > 0.01 ~ "under01",
                                   padj < 0.01 & padj > 0.001 ~ "under001",
                                   padj < 0.001 & padj >0 ~ "under0001",
                                   TRUE ~ "over01"))
      
      #threshold <- 8
      ggplot(results_sub_plot1 %>% as.data.frame(), aes(x=baseMean, y=log2FoldChange, color=color_sig, label=geneid)) +
          geom_point(size=1) +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange > threshold,],
          #           aes(x=baseMean, y=threshold), shape = 2, colour="#d83a00") +
          # geom_point(data= results_sub_plot1[results_sub_plot1$log2FoldChange < -threshold,],
          #           aes(x=baseMean, y=-threshold), shape = 2, colour="#d83a00") +
          #ylim(-threshold,threshold) +
          scale_x_log10() +
          scale_color_manual(values = c("gray30","#d83a00","#ff9b76","#ffd4c4")) +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange > 0),
                  nudge_y = 20,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          geom_text_repel(data = subset(results_sub_plot1, color_sig=="under0001" & baseMean > 100 & log2FoldChange < 0),
                  nudge_y = -20,
                  #nudge_x=-10,
                  #force_pull   = 10,
                  force         = 65,
                  max.overlaps = Inf,
                  box.padding = 0.5,
                  segment.color = "grey50",
                  direction     = "both") +
          ylab(paste0("log2 Fold Change ",celltypes , conditions[2]," vs ",conditions[1] ," RNA")) +
          ggtitle(paste0("Results ",celltypes , conditions[2]," vs ",conditions[1] )) +
          theme_bw() 
      
      
      ## Export files
      
      # write.table(dds_sub_counts,
      # file = paste0(workingdir,outdir,suboutdir2,"CountsNormalized_",celltypes,days[1],days[2],".txt"),
      #     quote = FALSE, row.names = TRUE)
      # write.csv(vsd_sub_data,
      #     paste0(workingdir,outdir,suboutdir2,"VSData_",celltypes,days[1],days[2],".csv"),
      #     quote = FALSE)
      # write.table(results_sub,
      #     file = paste0(workingdir,outdir,suboutdir2,"Results_DESeq_",celltypes,days[1],days[2],".txt"),
      #     quote = FALSE, row.names = TRUE)

      results_return <- results_sub %>% as.data.frame() %>% rownames_to_column("Geneid")
      results_return$Comparison <- paste0("Comp_",celltypes,days[1],days[2])
      #results_return

#   })
# }) 


```



```{r}
sessionInfo()
```