---
title: "R_quick_plot genes"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RNA analysis
Gene plotting with output from previous script.

```{r message=FALSE}

rm(list=ls())

library(RColorBrewer)
library(tidyverse)
library(ComplexHeatmap)


```




### Set dirs
```{r}

workingdir="~/Dropbox (The Francis Crick)/DP_cisReg/"
subworkinput="outputs_RNApilot_1/"


```


## Load RNA data

Normalised RNA counts

```{r }


dds_counts <- read.table(paste0(workingdir,subworkinput,"featurecounts.normCounts.txt"),stringsAsFactors =FALSE)


```




## Colors and shapes
Before more plotting, let's get some metadata organised

```{r }
sorted_gate <- c("pMN","DP","p3","neur")
sorted_conditions <- c("500","UPSAG","dRA2UPSAG","dRA")


shapes4_manual = c(18,15,16,17) # these are block
shapes5_manual = c(25,21,22,23,24) # these are filled
shapes4_fill_manual = c(23,21,22,24)
# 
# # for Days
# red_colors <- c("#fadede","#f3aaaa","#e96666","#cf1e1e")


#color_gates <- c("#b30000","#800080","#009640","#696969")
#color_gates <- c("#b30000","#9a009a","#005ab3","#696969")
color_gates <- c("#e60000","#cd00cd","#0073e6","#696969")

# for Days
colors_greys <- c("#f6f6f6","#808080","#333333")

# conditions
#colors_conditions <- c("#b81bb8","#b81b1b","#1bb81b","#1bb8b8")
colors_conditions <- c("#e67300","#4d9a00","#cdcd00","#0073e6")

```


## Genes to look at

```{r}

dds_counts_plot <- dds_counts %>% 
  as.data.frame() %>%
  rownames_to_column("geneid") %>%
  gather(sampleid, counts_norm, starts_with("D")) %>%
  separate(sampleid,into=c("Day","Condition","Gate","Rep"), sep="_", remove=FALSE) %>%
  mutate(DayGate=factor(paste(Day,Gate,sep="_")),
         DayCondition=paste(Day,Condition),
         Gate=factor(Gate, levels=sorted_gate),
         Condition=factor(Condition, levels=sorted_conditions))

```

### Basic QC

```{r fig.height=9, fig.width=9}

geneOI <- c("Sox2",
            "Dbx1","Irx3","Pax6","Nkx6-1","Olig2","Nkx2-2","Foxa2","Shh","Arx",
            "Tubb3","Sim1","Mnx1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```


### Temporal genes from pilot analysis

p3 d6 to d7

```{r fig.height=6, fig.width=9}

geneOI <- c("Lin28a","Nfib","Rbp1","H19","Fabp7","Hoxb6","Nkx6-2","Phox2b")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI) %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Condition,y=counts_norm)) +
  stat_summary(aes(fill=Day),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Day), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  #scale_fill_manual(values=color_gates) +
  #scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Gate, scales = "free_y") +
  theme_bw()

```

### What are those neurons?

Just D7:

- dRA: Fev and Gata3 point to more anterior (5HT generation?)
- dRA also high in Pou3f1 (Alcam), expressed in spinal accessory column (Cervical level)

```{r fig.height=6, fig.width=10}

geneOI <- c("Tubb3","Mnx1","Isl2","Isl1","Neurog2","Lhx3","Lhx1",
            "Slc18a3","Irx3",
            "Nkx2-2","Sim1","Neurog3","Phox2b","Pou3f1","Alcam",
            "Abca1","Fgf10",
            "Gata3","Fev","Slc17a8","Cyp26b1",
            "Pou4f2","Lpar1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Day=="D7") %>% mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_wrap(~ geneid, scales = "free_y", ncol=4) +
  theme_bw()

```

### Hoxes

```{r fig.height=6, fig.width=9}

#geneOI <- grep("Hoxc", dds_counts_plot$geneid, value = TRUE)

geneOI <- c("Hoxa1","Hoxa3","Hoxa5","Hoxb2","Hoxb5","Hoxb7","Hoxb8","Hoxc4","Hoxc5","Hoxc6")

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Day=="D7"), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_wrap(~ geneid, scales = "free_y", ncol=3) +
  theme_bw()

```

### What are DPs

```{r fig.height=9, fig.width=9}

geneOI <- c("Dll1","Dll3","Neurod4","Nkx2-9",
            "Hey1","Hey2","Neurog1","Neurog2","Hes1","Hes4","Notch1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>% 
        mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Condition),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Condition), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=colors_conditions) +
  scale_color_manual(values=colors_conditions) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Day, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         filter(Gate !="neur") %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Day),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Day), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  #scale_fill_manual(values=color_gates) +
  #scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```

Genes from Rayon et al

```{r fig.height=9, fig.width=9}

geneOI <- c("Olig2","Olig1","Nkx2-2","Nkx2-9","Nkx6-1","Nkx6-2","Fabp7",
            "Sox9","Neurod4","Neurog1","Neurog2","Neurog3")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         filter(Gate !="neur") %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Day),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Day), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  #scale_fill_manual(values=color_gates) +
  #scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```


Rayon's "gliogenic score"

FABP7, SOX9, SOX10, PDGFRA, CSPG4, FGFR3, FGFBP3, DBI, SLC1A3, HOPX, ALDH1L1

```{r fig.height=9, fig.width=9}

geneOI <- c("Fabp7","Sox9","Sox10","Pdgfra","Cspg4","Fgfr3","Fgfbp3","Dbi","Slc1a3","Hopx","Aldh1l1")


ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Day,y=counts_norm)) +
  stat_summary(aes(fill=Gate),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Gate), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  scale_fill_manual(values=color_gates) +
  scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

ggplot(dds_counts_plot %>% filter(geneid %in% geneOI & Condition %in% c("500","UPSAG","dRA2UPSAG")) %>%
         filter(Gate !="neur") %>%
         mutate(geneid=factor(geneid, levels=geneOI)), 
       aes(x=Gate,y=counts_norm)) +
  stat_summary(aes(fill=Day),
    fun = mean, geom="bar", alpha=0.9, width=0.7,position=position_dodge(0.7)) +
  geom_point(aes(fill=Day), alpha=0.6, position = position_dodge(width = 0.7),color="black") +
  #geom_col(position="dodge",aes(fill=DayGate)) +
  #scale_fill_manual(values=color_gates) +
  #scale_color_manual(values=color_gates) +
  scale_shape_manual(values=shapes4_fill_manual) +
  facet_grid(geneid ~ Condition, scales = "free_y") +
  theme_bw()

```



```{r}
sessionInfo()
```


